Stage-Splitting a Modal Language

### 2014-10-08 paper/relatedwork.tex
more notes on metaprogramming

Davies ...

Taha's approach is similar to Davies, but extends Davies's approach
with ``cross-stage persistency'' which allows a variable to be used in
any later stage and a ``run'' construct that allows evaluating an
expression.  Our language constructs therefore are also close to
staged programming such as proposed by Taha.  In fact our static
semantics is essentially identical to Taha's.  Though, as Davies, we
don't include ``run'' in the language.  The dynamic semantics of
staged languages... COMPLETE

The key difference between our work and Taha's and similar work on
staging is our splitting algorithm that can translate the staged code
to a different code that can run significantly (sometimes
asymptotically) faster. COMPARE DYNAMIC SEMANTICS

 Unlike the evaluati

Multi-stage, a.k.a., meta programming or staged programming, gives the
programmer linguistic facilities consisting of language constructs
(usually also a type system, including a dynamic semantics) for
expressing programs where inputs may come in stages.  Earlier work on
staged programming was motivated by understanding the effects of
binding-time analysis that partial evaluation relied on and allowing
control over staging so that the programmer can control the
termination and efficiency characteristics of staged programs.
Several different approaches to staged programming has been proposed.

>>>>

There's another staging technique known as {\em metaprogramming} (\cite{taha-thesis-99}, \cite{devito13}, \cite{davies01}).
The differences between it and partial evaluation can be subtle.
The core idea of metaprogramming is to interpret stage \bbtwo\ values as {\em code} in some object language
which stage \bbone\ can manipulate.
This has three implications:
\begin{itemize}
\item Stage annotations are now mandatory, rather than optional.
\item The $\next$-contents-duplicating semantics is now intended behavior [this needs to be verified].
This changes the cost of terms from what they would be in \lang's semantics, 
but that's merely something for the programmer to keep in mind.
\item In metaprogramming systems, it now makes sense to define an operator, \texttt{run}, 
which evaluates stage \bbtwo\ code to get an answer back at stage \bbone.
When you think about it, this is a form a backwards information flow.
For this reason, the \texttt{run} feature is in direct conflict with stage-splitting,
which by its definition cannot work for a language that allows backwards control flow.
\end{itemize}
